---
title: "GTM Boat Use"
author: "SDK"
format: revealjs
editor_options: 
  chunk_output_type: console
---
```{r}
#| include: false


library(here)
library(tidyverse)
library(lubridate)
library(janitor)
library(readxl)
library(kableExtra)
```

```{r}
#| include: false


df <- readr::read_csv(here('data', '2023-10_floatplan_responses.csv')) %>% janitor::clean_names()
```

```{r}
#| include: false


df2 <-
df %>% 
  select(date_you_plan_to_launch, vessel_details, who_is_the_primary_captain) %>% 
  rename(date = date_you_plan_to_launch,
         vessel = vessel_details,
         captain = who_is_the_primary_captain) %>% 
  mutate(date = lubridate::as_date(date, format = "%m/%d/%Y"))

```

```{r}
op <- readxl::read_xlsx(here('data', '2023-10_operators.xlsx'), sheet = "operators2") %>% janitor::clean_names()

op2 <- 
op %>% 
  rename(date = date_you_plan_to_launch) %>% 
  mutate(date = lubridate::as_date(date)) %>%
  tidyr::pivot_longer(cols = 2:4,
                           names_to = 'operatorno', values_to = 'name' 
                           ) %>% 
  filter(!is.na(name)) %>% select(-operatorno) %>%
  distinct()

other <- readxl::read_xlsx(here('data', '2023-10_operators.xlsx'), sheet = "others2") %>% janitor::clean_names()

other2 <-
  other %>% 
  rename(date = date_you_plan_to_launch) %>% 
  mutate(date = lubridate::as_date(date)) %>% 
  tidyr::pivot_longer(cols = 2:5,
                      names_to = 'indivno',
                      values_to = 'name') %>% 
  filter(!is.na(name)) %>% select(-indivno) %>% 
  distinct()



cap2 <- df2 %>% select(date, captain) %>% distinct() %>% rename(name = captain)

names <- bind_rows(op2, other2, cap2) %>% distinct()

rm(op, other, op2, cap2, other2)

```

This summarizes data from the GTMNERR online Float Plan responses. 

-   The online Float Plan has been used as the primary Float Plan since 2018

-   The data was downloaded on 2023-11-02

# Which boat(s) are used the most?

## The UNF boat is used the most

```{r}
df2 %>% 
  filter(!is.na(vessel)) %>% 
  filter(!grepl("Canoe", vessel)) %>% 
  filter(!grepl("Kayak", vessel)) %>% 
  group_by(vessel) %>%
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  ggplot(aes(x = fct_reorder(vessel, -count, max), y = count)) +
  geom_col() +
  geom_text(aes(label = count), hjust = -1, color = "black") +
  ylim(0,450) +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "", 
       y = "Number of Days in Use")
```

# Boating Days for Captains*

*When an individual is specifically listed as the primary captain for the day.

## Number of Boat Days Per Captain

```{r}
df2 %>% 
  group_by(captain) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  ggplot(aes(x = fct_reorder(captain, -count, max), y = count)) +
  geom_col() +
  geom_text(aes(label = count), hjust = -1, color = "black") +
  ylim(0,400) +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "", 
       y = "Number of Boat Days")
```

## Number of Boat Days Per Captain (this year)

```{r}
df2 %>% 
  filter(date >= as_date("2023-01-01")) %>% 
  group_by(captain) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  ggplot(aes(x = fct_reorder(captain, -count, max), y = count)) +
  geom_col() +
  geom_text(aes(label = count), hjust = -1, color = "black") +
  ylim(0,60) +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "", 
       y = "Number of Boat Days")
```

## Frequency of Boating per Captain (since 2018)

```{r}
df2 %>% 
  filter(date >= as_date("2018-01-01")) %>% 
  filter(!grepl("Canoe", vessel)) %>% 
  filter(!grepl("Kayak", vessel)) %>% 
  filter(!grepl("Mike Dickson", captain)) %>%
  filter(!grepl("Scott Eastman", captain)) %>%
  filter(!grepl("Rauney Mitchell", captain)) %>%
  ggplot(aes(x = date, y = captain)) +
  geom_point() + 
  scale_x_date(date_breaks = "year", date_labels = "%Y", minor_breaks = "1 month", position = "top") +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "",
       y = "")
```

## Frequency of Boating per Captain (this year)

```{r}
df2 %>% 
  filter(date >= as_date("2023-01-01")) %>% 
  filter(!grepl("Canoe", vessel)) %>% 
  filter(!grepl("Kayak", vessel)) %>% 
  filter(!grepl("Mike Dickson", captain)) %>%
  filter(!grepl("Scott Eastman", captain)) %>%
  filter(!grepl("Rauney Mitchell", captain)) %>%
  ggplot(aes(x = date, y = captain)) +
  geom_point() + 
  scale_x_date(date_breaks = "month", date_labels = "%b", position = "top") +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "",
       y = "")
```

# Boating Days all staff

This includes staff as well as volunteers that have been certified to operate vessels.

## Number of Boat Days Per Staff (since 2018)

```{r}
names %>% 
  group_by(name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  ggplot(aes(x = fct_reorder(name, -count, max), y = count)) +
  geom_col() +
  geom_text(aes(label = count), hjust = -1, color = "black") +
  ylim(0,400) +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "", 
       y = "Number of Boat Days")
```

## Number of Boat Days Per Staff (this year)

```{r}
names %>% 
  filter(date >= as_date("2023-01-01")) %>% 
  group_by(name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  ggplot(aes(x = fct_reorder(name, -count, max), y = count)) +
  geom_col() +
  geom_text(aes(label = count), hjust = -1, color = "black") +
  ylim(0,50) +
  coord_flip() +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "", 
       y = "Number of Boat Days")
```

## Frequency of Boating per staff (this year)

```{r}
names %>% 
  filter(date >= as_date("2023-01-01")) %>% 
  ggplot(aes(x = date, y = name)) +
  geom_point() + 
  scale_x_date(date_breaks = "month", date_labels = "%b", position = "top") +
  theme_bw() +
  theme(axis.text = element_text(size = 12, color = "black")) +
  labs(x = "",
       y = "")
```